---
title: "Weight Lift Quality Prediction"
date: "Sunday, January 25, 2015"
output: html_document
---
Background

In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the quality of their weight lifting exercise. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Load caret library
```{r,echo=F}
library(caret)
```
1. Data preparation
read in training and testing raw data.
```{r}
setwd("C:\\")
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```
remove variables with no value, sparse values, and non-relavant variables.
```{r}
training1 <- training[,!is.na(training[1,])]
rmlist <- c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp",
            "new_window","num_window","kurtosis_roll_belt","kurtosis_picth_belt","kurtosis_yaw_belt",
"skewness_roll_belt","skewness_roll_belt.1","skewness_yaw_belt","max_yaw_belt","min_yaw_belt",
"amplitude_yaw_belt","kurtosis_roll_arm","kurtosis_picth_arm","kurtosis_yaw_arm",
"skewness_roll_arm","skewness_pitch_arm","skewness_yaw_arm","kurtosis_roll_dumbbell",
"kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell",
"skewness_yaw_dumbbell","max_yaw_dumbbell","min_yaw_dumbbell","amplitude_yaw_dumbbell",
"kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm","skewness_roll_forearm",
"skewness_pitch_forearm","skewness_yaw_forearm","max_yaw_forearm","min_yaw_forearm",
"amplitude_yaw_forearm")
training1 <- training1[,!(names(training1) %in% rmlist)]
```
remove varibale with no variance and reduce pair-wise correlations with correlation threshold of 0.8. 
This will prevent buiding a overfitting model.
```{r}
training2 <- training1[,-nearZeroVar(training1)]
training2 <-  training1[,-findCorrelation(cor(training1[,c(1:52)]), .8)]
```
Now, only 40 predictors are left to build a model
```{r}
names(training2)
```

2. Model Training
Cross-validation dataset preparation: split training2 into cross-validation training and testing sets
```{r}
set.seed(1)
inTrainCV <- createDataPartition(training2$classe, p = 0.7, list=FALSE)
trainingCV <- training2[inTrainCV,]
testingCV <- training2[-inTrainCV,]
```

setup trainControl with 10-fold cross-validation
```{r}
fitControl <- trainControl(method = "repeatedcv",number = 10,repeats = 10)
```

We first tried random-forest method, however, it is prohibitively time consumeing, therefore, we turned to quadratic discriminant analyis method to train model with the option of standarizing predictors.
```{r}
set.seed(2)
modFit <- train(classe~., method = "qda", data=trainingCV, trControl = fitControl, 
                preProcess = c("center","scale"))
modFit
```
As shown above, we achived accuracy of 0.857 on traingCV dataset, we expect the out-of-sample error to be larger for testingCV dataset 

3. Cross-validation and out-of-sample error

use the model to predict the response on cross-validation testing dataset - testingCV 
```{r}
predtestingCV <- predict(modFit,testingCV)
conMatCV <- confusionMatrix(predtestingCV,testingCV$classe)
conMatCV$table
```

calculate accuarcy 
```{r}
testingCV_accurary <- sum((predtestingCV==testingCV$classe))/dim(testingCV)[1]
testingCV_accurary 
```
The accuracy from testingCV is indeed very close to that from the trainingCV, the out-of-sample error is 
```{r}
1-testingCV_accurary
```

4. Prediction on testing dataset
```{r}
predict_testing <- predict(modFit,testing)
predict_testing
```

5. Generate prediction to each of the testing case in testing dataset
```{r}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(predict_testing)
```
19 out of 20 predictions are correct

6. Conlcusion
In summary, due to time constrain, we used a fast method, quadratic discriminant analysis, to build a model to predict weight lifting quality from five difference types of sensor data. The Acuracy of this model is about 86%, which was verified by cross-validation.  
